---
title: "Impact of CO2 Emission on Agricultural Production"
author: "Justin Chow and Max Payen"
date: "2023-03-19"
output: html_document
---
```{r setup, echo=FALSE, message=F, warning=F, include=FALSE}

#==============================================================================
#   This chunk will be used for every assignment
#==============================================================================

  # Clear the working space
    rm(list = ls())
  
  # Set working directory
  setwd("C:/Users/14087/Documents/master/econ 2509/project")

### Load the packages (all must have been installed)
    library(tidyverse)
    library(doBy)
    library(foreign)
    library(ivpack)
    library(knitr)
    library(lmtest)
    library(readstata13)
    library(sandwich)
    library(stargazer)
    library(AER)
    library(gdata)
    library(wooldridge)
    library(openintro)
    library(remotes)


cse=function(reg) {
        rob=sqrt(diag(vcovHC(reg, type="HC1")))
        return(rob)
        }

ivse=function(reg) {
			rob=robust.se(reg)[, 2]
			return(rob)
      }
```

```{r data, echo=FALSE, warning=FALSE, include=FALSE}

# load agrico2.csv dataset
agrico2=read.csv("agrico2.csv", header=T, sep=",")
# reading data and the summary data
str(agrico2)
# handle missing data
agrico2 <- na.omit(agrico2)

```

## Descriptive Statistics

```{r stargazer, echo=FALSE, warning=FALSE, comment="" }

# Descriptive Statistics
stargazer(agrico2[c("co2_emission", "land_area", "food_index", "lv_stress")], type="text", digits=2, summary.stat=c("n", "mean", "median", "sd", "min", "max"), title="Descriptive Statistics")

```

The Agriculture data set contains information on 60 countries from 2019 with the level of CO2 emissions, percent of land area used for agriculture, food production index, and level of freshwater stress. The median for the level of CO2 emission, food production index, and level of freshwater stress is lower than their mean, which suggests that all of the distributions are negatively skewed. The median for the percent of land area used for agriculture is higher than its mean, which suggests that the distribution is positively skewed.

## Data Analysis and Visualizations

```{r image1, echo=FALSE, warning=FALSE, message=FALSE, comment=""}

# Line plot
ggplot(agrico2, aes(x=log(co2_emission), y=crop_index)) + 
  geom_point(col="blue") + 
  geom_text(aes(label=code), hjust=1, vjust=1) + 
  labs(title = "Crop production index on Level of CO2 emissions", x = "Level of CO2 emissions", y = "Crop production index") +
  stat_smooth(method = "lm", col = "red", se=FALSE)

```

There seems to be a negative relationship between the level of CO2 emissions and the crop production index of all 60 countries. There are two European country outliers - Croatia (HRV) and Armenia (ARM) - that have much smaller crop production index compared to the rest of the countries, but it does not seem to effect the negative linear relationship between the level of CO2 emission and crop production index.

```{r image2, echo=FALSE, message=F, warning=FALSE, comment=""}

# Bar plot
ggplot(agrico2, aes(x=crop_index, y=reorder(code, -crop_index), fill = co2_emission)) +
  geom_bar(stat="identity") +
  xlab("Crop Production Index") +
  ylab("Country Code") +
  theme_minimal(base_size = 10)

```

Developed countries such as Australia, Germany, and Poland have lower crop production index and higher levels of CO2 emissions. Developing countries such as Romania, Nicaragua, and Honduras have higher crop production index and lower levels of CO2 emissions. 

These trends suggest that a larger country will have a lower crop production index and higher levels of CO2 emissions due to the increase in emissions reducing the number of crops produced.

```{r hist, echo=FALSE, message=F, warning=FALSE, comment=""}

# Categorical variable named lowcrop
agrico2$lowcrop[agrico2$crop_index <= 103.41] <- "Low value" 
agrico2$lowcrop[agrico2$crop_index > 103.41] <- "Higher value"

# Histogram plot
ggplot(agrico2, aes(x = co2_emission, fill = lowcrop)) + 
  geom_histogram(binwidth=0.25, position="identity", alpha=0.4)

```

The countries with high or low crop production indexes have distributions that are negatively skewed since the median is greater than the mean level of CO2 emissions. The higher crop production indexes are located near lower levels of CO2 emissions, while a few countries have high crop production indexes and high levels of CO2 emissions. These outliers represent countries that rely on machinery and factory production when processing their crops to be sold.

```{r relation, echo=FALSE, message=F, warning=FALSE, comment=""}

# transform GDP growth rate from character to numeric type
agrico2$GDP = as.numeric(agrico2$GDP) 
agrico2$GDP_growth = as.numeric(agrico2$GDP_growth) 

# Relationship between GDP per capita growth rate and CO2 Emission 
ggplot(agrico2, aes(x = GDP_growth, y=co2_emission, size = GDP, colour = code)) +
  geom_point() +
  theme(legend.position = "right") + 
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  scale_x_continuous(limits = c(-2, 8)) +
  labs(title = 'GDP per capita growth rate and CO2 Emission', subtitle = 'Year: 2019', x = 'GDP per capita growth rate', y = 'Level of CO2 Emissions')

```

The relationship between GDP per capita growth rate and level of CO2 emissions is slightly negative relative to the size of the GDP. Specifically, the USA is an outlier in this relationship due to its higher level of CO2 emissions but lower GDP per capita growth rate. The size of the GDP shows that larger countries such as the USA, India, and China have the highest GDP. As a result, they all have higher levels of CO2 emission but the GDP per capita growth rate is relatively smaller than the rest of the countries. 


## Regression Analysis

```{r regression, echo=FALSE, message=F, warning=FALSE, comment=""}

# transform type of GDP variables from character to numeric type
agrico2$GDP_per_capita = as.numeric(as.character(agrico2$GDP_per_capita))

# logarithm of co2_emission
agrico2$lco2 <- log(agrico2$co2_emission)
# logarithm of GDP (per capita)
agrico2$l_GDP <- log(agrico2$GDP_per_capita)
# logarithm of lv_stress
agrico2$lstress <- log(agrico2$lv_stress)

# Regressions
# Run 4 regression models using lm command
regr1<-lm(crop_index~lco2, data=agrico2)
regr2<-lm(crop_index~lco2+food_index, data=agrico2)
regr3<-lm(crop_index~lco2+food_index+land_area, data=agrico2)
regr4<-lm(crop_index~lco2+food_index+land_area+lstress, data=agrico2)
regr5<-lm(crop_index~lco2+food_index+land_area+lstress+l_GDP, data=agrico2)
regr6<-lm(crop_index~lco2+food_index+land_area+lstress+l_GDP+I(lco2*lstress), data=agrico2)

# Using stargazer to display your regression results 
stargazer(regr1, regr2, regr3, regr4, regr5, regr6,
          se=list(cse(regr1), cse(regr2), cse(regr3), cse(regr4), cse(regr5), cse(regr6)), 
          title="Level of CO2 emissions on Crop Production Index", type="text", 
          df=FALSE, digits=3, omit.stat=c( "f"))

```

In regression 1, an increase in a country's level of CO2 emission by 1% reduces its crop production index by 0.0315. With an absolute t-statistic of 2.00, the level of CO2 emission is statistically significant at the 5% significance level.

In regression 2, holding the food production index constant, an increase in a country's level of CO2 emission by 1% reduces their crop production index by 0.0199. With an absolute t-statistic of 1.68, the level of CO2 emission is statistically significant at the 10% significance level. Since the estimate of the co2_emission variable increases from regression 1 to 2, we can conclude that regression 1 suffers from downward omitted variable bias. 

In regression 3, the control variable agricultural land has a standardized beta coefficient of -0.091. Holding the co2_emission and food production index constant, an increase in agricultural land by 1% decreases the crop production index by 0.091. This control variable also changes the effect of a country's levels of CO2 emission on the crop production index. Holding everything else constant, an increase in a country's level of CO2 emission by 1% reduces the crop production index by 0.0191. With an adjusted R2 of 0.589, we can say that this model explains 58.9% of the variation in the crop production index, which is relatively high.

According to both regression 2 and 3, the level of CO2 emission in regression 3 has a slightly smaller negative effect on the crop production index. The addition of agricultural land slightly increases the level of CO2 emission beta coefficient from -1.99 to -1.91. This is a sign of downward omitted variable bias in regression 2.

In regression 4, the control variable logarithm of level of freshwater stress has a standardized beta coefficient of 0.34. Holding everything else constant, an increase in the level of freshwater stress by 1% increases the crop production index by 0.34. In this regression, the level of freshwater stress is not statistically at the 10% level. This explains the adjusted R2 of 0.583 meaning the model explains 58.3% of the variation in the crop production index, which is a slight drop from the previous model. Holding everything else constant, an increase in a country's level of CO2 emission by 1% reduces the crop production index by 0.02. 

According to both regressions 3 and 4, the level of CO2 emission has a larger negative effort on the crop production index. The addition of level in freshwater stress increases the level of CO2 emission beta coefficient from -1.91 to -2.00. This is a sign of downward omitted variable bias in regression 3. 

In regression 5, an increase in a country's level of CO2 emission by 1% reduces their crop production index by 0.0214, holding everything else constant. If a country has a higher GDP per capita, the level of CO2 emission reduces the crop production index by more. In this regression, GDP per capita is also not statistically at the 10% level, despite the variable of interest is statistically significant at the same level. This explains the adjusted R2 of 0.576 meaning the model explains 57.6% of the variation in the crop production index, which continues the drop from the previous model. Holding everything else constant, an increase in a country's level of CO2 emission by 1% reduces the crop production index by -0.0214.

In regression 6, the addition of interaction term between the level of freshwater stress and level of CO2 emissions drastically increases the negative impact of CO2 emissions on crop production index. Holding everything else constant, the effect of increasing the level of CO2 emission by 1% on the crop production index is -0.0446+0.0117*log(lv_stress)%. 

A country with a higher level of freshwater stress (or the proportion of total renewable freshwater resources) reduces the negative effect of CO2 emission on the crop production index. While the interaction term between the level of freshwater stress and level of CO2 emissions is slightly statistically insignificant at the 10% level, the model explained 57.9% of the variation in crop production index, which increased from the previous model.

In regression 6, any change in crop production index by an increase in level of CO2 emissions by 1% is economically significant, especially with the minimum crop production index reducing by 0.0446. Despite this, when looking at the regression of level of freshwater stress on crop production index is economically insignificant due to the control variable being statistically insignificant at the 10% level. The difference in crop production index comes from the food production index and percentage of agricultural land.    


## F-Test on Interaction Term

```{r f-test, echo=FALSE, message=F, warning=FALSE, comment=""}

# F-test on Interaction term between level of CO2 emission and level of freshwater stress from regression 6
lht(regr6,c("lco2=lstress"), white.adjust="hc1")

```

The F-test above suggests that the interaction term between the logarithm of CO2 emission and logarithm of freshwater stress is not statistically significant at a 10% significance level. This means we fail to reject the null and the coefficient of the logarithm of  is equal to 0.


## Predict for Crop Production Index

```{r predict, echo=FALSE, message=F, warning=FALSE, comment=""}

# Using the results of regression 3 and predict the crop production index for a country in the 75th percentile with log of CO2 emissions = 1.98, food production index = 109.31, and land_area = 51.40.
newdata=data.frame(lco2 = 1.98, food_index=109.31, land_area=51.40)
predict(regr3, newdata)

```

Using the best performing regression model (explaining 58.9% of the variation in the crop production index), a country with an increase in the level of CO2 emission around 2%, a food production index around 110, and the percentage of agricultural land slightly above 50% has a crop production index of 105. This crop production index was reduced from the 75th percentile to an index closer to the 50th percentile.   


## Probit- and Logit- Model

```{r pl, echo=FALSE, message=F, warning=FALSE, comment=""}

# Create dummy variable of crop index for countries above the 50th percentile 
agrico2$medcrop <- ifelse(agrico2$crop_index < 103.41, 1, 0)

# Run a probit- and logit- models using glm command
p1=glm(medcrop~lco2+food_index+land_area, family=binomial(link="probit"), x=TRUE, data=agrico2)
l1=glm(medcrop~lco2+food_index+land_area, family=binomial, x=TRUE, data=agrico2)

# regression table
stargazer(p1, l1, 
          se=list(NULL, NULL), 
          title="Probit- and Logit- Model of Medium-Sized Countries Crop Index",
          type="text", df=FALSE, digits=3,
          omit.stat=c( "f"))

```


## Marginal Error

```{r marginal, echo=FALSE, message=F, warning=FALSE, comment=""}

# calculate marginal error from probit- and logit- regression
library(erer)
fm1=maBina(p1, x.mean=FALSE, rev.dum=TRUE, digits=3)
fm2=maBina(l1, x.mean=FALSE, rev.dum=TRUE, digits=3)

# Using stargazer to display the marginal effects of the probit and logit regressions
stargazer(p1, fm1, l1, fm2,
          se=list(NULL, NULL, NULL, NULL), 
          title="Marginal Effects",
          type="text", df=FALSE, digits=3,
          keep.stat=c("n","ll"))

#calculate the pseudo-R2
pseudoR2p1=(p1$null.deviance-p1$deviance)/p1$null.deviance
round(pseudoR2p1, digits=3)
pseudoR2l1=(l1$null.deviance-l1$deviance)/l1$null.deviance
round(pseudoR2l1, digits=3)

```

In the probit model, the estimates suggest that everything else is the same, an increase in the level of CO2 emission by 1%, on average, will increase the probability of crop index for most of the countries, on average, 3.6% more than other countries. The effect is not statistically significant at more than a 10% significance level.

In the logit model, the estimates suggest that everything else is the same, an increase in the level of CO2 emission by 1%, on average, will increase the probability of a country within the vast majority of crop index by 2.9%. The effect is also not statistically significant at more than a 10% significance level.

